---
title: "Crime_Project"
author: "Chris"
date: "Wednesday December 2, 2015"
output: html_document
---

The goal is to plot gis points of crime in Denver, see patterns and predict the incidences of crime or see the clusters. To start I am not sure on the focus of the project, but I am leaning towards violent crime. At the end, I am going to try and see 

1)Initial EDA, just to see the basis of the data that I am looking at. 

```{r}
crime <-read.csv("C:/Users/BandC/Desktop/Predictive Model/crime.csv")
attach(crime)
summary(crime)
str(crime)
dim(crime)
names(crime)
str(INCIDENT_ID)
summary(INCIDENT_ID)#irrelevant
summary(OFFENSE_ID)
summary(OFFENSE_CODE)
summary(OFFENSE_CODE_EXTENSION)
summary(OFFENSE_TYPE_ID)#This might prove to be useful, the others are not thus far, table
summary(OFFENSE_CATEGORY_ID)#For crime, this will also prove to be useful,table 
summary(FIRST_OCCURENCE_DATE)
summary(REPORTED_DATE)# This will be useful to gather data on when they occured, table
summary(INCIDENT_ADDRESS)#Should look at top 5 to 10, maybe break it down by GIS
summary(GEO_X)#Irrelevant
summary(GEO_Y)#Irrelevant
summary(GEO_LON)#Irrelevant
summary(GEO_LAT)#Irrelevant
names(crime)
summary(DISTRICT_ID)#Put in table
summary(PRECINCT_ID)#Put in tables
summary(NEIGHBORHOOD_ID)#Will look at specific hoods for further investigation
summary(IS_CRIME)#irrelevant
summary(IS_TRAFFIC)#irrelevant


```

Before moving to plots, I want to look at some tables and possibly proportions to see if this gives any insight into what is most common and other factors to keep in mind.

```{r}
str(crime)
table(INCIDENT_ID)#This is not relevant, instead of crime ID, it is the chronological ID, which will probably not prove to be super helpful
table(OFFENSE_ID)
ID=head(OFFENSE_ID,5)
ID
prop.table(ID)#When looking at crime, instead of going for all crime, which might not be useful, due to the sheer volume of traffic tickets, jaywalking...etc. I might narrow it down later to look at just crimes against persons or property crime. I will re-evaluate the goal in a bit
table(OFFENSE_CODE)
prop.table(OFFENSE_CODE)
head(subset(crime[order(OFFENSE_CODE, decreasing=T),])[1:5]
?rank
names(crime)

sort(table(crime$OFFENSE_CODE), decreasing=TRUE)[1:5]# The top five most common
most_common=crime[which(OFFENSE_CODE==5441| OFFENSE_CODE==2999 | OFFENSE_CODE==5401 | OFFENSE_CODE==2305 | OFFENSE_CODE==2999),]
most_common
names(most_common)
most_commontwo <-subset(crime, OFFENSE_CODE==5441| OFFENSE_CODE==2999 | OFFENSE_CODE==5401 | OFFENSE_CODE==2305 | OFFENSE_CODE==2999, select=c(DISTRICT_ID, PRECINCT_ID))
most_commontwo
head(most_commontwo)
table(most_commontwo)
prop.table(most_commontwo)
names(crime)
head(crime$OFFENSE_CODE)
table(crime$OFFENSE_CODE)
head(crime$OFFENSE_CODE, des)
code=crime[order(OFFENSE_CODE),]
code
sort(table(OFFENSE_CODE),decreasing=TRUE)[1:3]
head(OFFENSE_CODE_EXTENSION)
tail(OFFENSE_CODE_EXTENSION)
summary(OFFENSE_CODE_EXTENSION)#PROBABLY NOT VERY USEFUL
head(OFFENSE_TYPE_ID)
tail(OFFENSE_TYPE_ID)
table(OFFENSE_TYPE_ID)
type=crime[order(OFFENSE_TYPE_ID),]

type
sort(table(OFFENSE_TYPE_ID),decreasing=TRUE)[1:3]

head(OFFENSE_CATEGORY_ID)
tail(OFFENSE_CATEGORY_ID)
table(OFFENSE_CATEGORY_ID)#This is a good breakdown of the crimes, violent, auto..etc. subsetting these will probably prove to be useful for graphing and predicting
head(FIRST_OCCURRENCE_DATE)
head(LAST_OCCURRENCE_DATE)#Not sure the significance of these two, will probably drop these rows, as they are not the reported date
names(crime)
head(REPORTED_DATE)
#MIGHT WANT TO CHANGE DATE AND LEAVE OUT TIME (HOUR) OR CREATE A NEW COLUMN OR TWO WITH DATE ONLY AND MONTH ONLY
table(REPORTED_DATE)
day=sort(table(REPORTED_DATE),decreasing=TRUE)[1:5]
day
names(crime)
head(INCIDENT_ADDRESS)
sort(table(INCIDENT_ADDRESS),decreasing=TRUE)[1:3]#Looks like no address given is most common, but maybe it will be given a lat and long
head(GEO_X)
sort(table(GEO_X),decreasing=TRUE)[1:3]#Still common with no X
head(GEO_LON)
sort(table(GEO_LON),decreasing=TRUE)[1:3]#LONGITUDE GIVEN 
sort(table(GEO_LAT),decreasing=TRUE)[1:3]#LAT AND LON MATCH WITH TABLE HAVE SAME NUMBER OTHER THAN COORDINATE, 39.73, -104.98, WILL PLOT ON GEO MAP AND SEE IF CAN FIND MATCHING ADDRESS
head(DISTRICT_ID)
table(DISTRICT_ID)#This will be useful to breakdown district and what is going and then subset with type of crime
head(PRECINCT_ID)
table(PRECINCT_ID)#More precincts than districts, will need to match a precinct with district, 32 precincts will not be an exact match to 7 districts, so some will cross-over
unique(crime$PRECINCT_ID)
head(NEIGHBORHOOD_ID)
table(NEIGHBORHOOD_ID)
sort(table(NEIGHBORHOOD_ID),decreasing=TRUE)[1:10]#Shows the neighborhoods with the most reported crime 
head(IS_CRIME)
table(IS_CRIME)#It is coded as 0 and 1, nothing else
head(IS_TRAFFIC)#It looks like Is_CRIME and IS_TRAFFIC are inversly related, if one is a 0 the other is 1. Will probably get rid of IS_TRAFFIC because traffic indcidents are not of interest
```{r}

```
The above was not necessary for every variable, but I like to do it before graphing to see if anything can be determined from the data and it looks like some will be very valuable where others intersect with other variables and are essentially duplicates. 

Graphing: Initial Graphs, GIS graphs and more intuitive graphs to follow

```{r}
library(ggplot2)
str(crime)
hist(crime$INCIDENT_ID)
hist(OFFENSE_CODE)
qplot(day,binwidth=30)#Will need to adjust binwidth
names(crime)
?frequency
require(graphics)
frequency(REPORTED_DATE)#Will try a frequency plot in another format
time(REPORTED_DATE)
qplot(OFFENSE_TYPE_ID)
qplot(OFFENSE_CATEGORY_ID)
plot(OFFENSE_CATEGORY_ID)
ADDY=sort(table(INCIDENT_ADDRESS),decreasing=TRUE)[1:3]
qplot(ADDY)#Will need to change up
qplot(DISTRICT_ID)
qplot(PRECINCT_ID)
qplot(NEIGHBORHOOD_ID)
qplot(OFFENSE_TYPE_ID, DISTRICT_ID)
qplot(DISTRICT_ID, OFFENSE_TYPE_ID, xlab="District",ylab="Type", main="Type by District", col=IS_CRIME)#Will need to narrow this down to be useful


```
In order to narrow down violent crime and crime against persons, I am going to create a new data frame that only has that information. 

```{r}
head(crime)
table(OFFENSE_CATEGORY_ID)#I am going to use aggravated-assault, other-crimes-against-persons, murder, robbery and sexual assualt
newdata <-subset(crime, OFFENSE_CATEGORY_ID=="aggravated-assault"|OFFENSE_CATEGORY_ID=="other-crimes-against-persons"|OFFENSE_CATEGORY_ID=="murder"|OFFENSE_CATEGORY_ID=="robbery"|OFFENSE_CATEGORY_ID=="sexual-assault")

head(newdata)
str(newcrimedata)#The total did go down a considerable amount, but I knew this since I am only interested in violent crime
dim(newdata)


```
With the focus clear on violent crime, I want to look at this new data set by looking at graphs and structure. 
```{r}
attach(newdata)
names(newdata)
table(newdata$OFFENSE_CATEGORY_ID)
length(newdata$OFFENSE_CATEGORY_ID)
table(newdata$OFFENSE_CATEGORY_ID)/length(newdata$OFFENSE_CATEGORY_ID)
library(ggplot2)
qplot(newdata$OFFENSE_CATEGORY_ID)#Other-crimes-against is the head
table(DISTRICT_ID)
hist(DISTRICT_ID, col=2)
qplot(NEIGHBORHOOD_ID, col=3, breaks=12,zoom=10)
table(IS_CRIME)
pairs(newdata)#Just to see if they are correlated in any way, not helpful
names(newdata)
qplot(DISTRICT_ID, PRECINCT_ID, data = newdata, col = OFFENSE_CATEGORY_ID,
      main = "DISTRICT", xlab="DIST_ID",
      ylab='PRECINCT')

qplot(DISTRICT_ID, data = newdata, col = OFFENSE_CATEGORY_ID,
      main = "DISTRICT", xlab="DIST_ID",
      )#sHOWS THE BREAKDOWN FOR EACH DISTRICT, PRETTY GOOD INFO
summary(OFFENSE_CATEGORY_ID)
qplot(PRECINCT_ID, data = newdata, col = OFFENSE_CATEGORY_ID,
      main = "CAT/DIST", xlab="CAT_ID",
     )
table(DISTRICT_ID)#Disrict  has very little crime

par=par(mfrow=c(4,3))
qplot(OFFENSE_CATEGORY_ID[DISTRICT_ID == 1], xlab="CATEGORY",
      main = "DISTRICT 1", col="IS_CRIME")
qplot(OFFENSE_CATEGORY_ID[DISTRICT_ID == 2], xlab="CATEGORY",
      main = "DISTRICT 2", col="IS_CRIME")
qplot(OFFENSE_CATEGORY_ID[DISTRICT_ID == 3], xlab="CATEGORY",
      main = "DISTRICT 3", col="IS_CRIME")
qplot(OFFENSE_CATEGORY_ID[DISTRICT_ID == 4], xlab="CATEGORY",
      main = "DISTRICT 4", col="IS_CRIME")
qplot(OFFENSE_CATEGORY_ID[DISTRICT_ID == 5], xlab="CATEGORY",
      main = "DISTRICT 5", col="IS_CRIME")
qplot(OFFENSE_CATEGORY_ID[DISTRICT_ID == 6], xlab="CATEGORY",
      main = "DISTRICT 6", col="IS_CRIME")
qplot(OFFENSE_CATEGORY_ID[DISTRICT_ID == 7], xlab="CATEGORY",
      main = "DISTRICT 7", col="IS_CRIME")#These show essentially what the table said but it is a good visual and shows how they change over districts and how the proportion is within a district




```
I want to try and look at times in a frequency table and see the days with most crime

```{r}
names(newdata)
newdata$REPORTED_DATE = as.Date(newdata$REPORTED_DATE, '%m/%d/%Y')

#plot date
qplot(newdata$REPORTED_DATE, binwidth=10)#peak in summer months
dayTWO=sort(table(newdata$REPORTED_DATE),decreasing=TRUE)[1:30]
dayTWO

qplot(newdata$REPORTED_DATE, binwidth=.1)




```
We can try to plot this also on a map of Denver since we have long and lat. I am going to break each district into a map to see where they are located in Denver. 
```{r}
names(newdata)
library(ggmap)

qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver
Denver + geom_point(data=newdata, aes(x=GEO_LON, y=GEO_LAT), 
                              color="dark green", alpha=.15, size=2)

newdata1 <-subset(crime, OFFENSE_CATEGORY_ID=="aggravated-assault"|OFFENSE_CATEGORY_ID=="other-crimes-against-persons"|OFFENSE_CATEGORY_ID=="murder"|OFFENSE_CATEGORY_ID=="robbery"|OFFENSE_CATEGORY_ID=="sexual-assault"&DISTRICT_ID=1)
newdatafirst<-subset(crime, DISTRICT_ID=1)
summary(newdatafirst)
str(newdatafirst)
dim(newdatafirst)
Denver + geom_point(data=newdata1, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=2)#subset not working properly
districtone<-subset(crime, DISTRICT_ID==1)
str(districtone)
library(ggmap)
Denver + geom_point(data=districtone, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=5)
districtwo<-subset(crime, DISTRICT_ID==2)
str(districtwo)

Denver + geom_point(data=districtwo, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=5)
districthree<-subset(crime, DISTRICT_ID==3)
str(districthree)
library(ggmap)
Denver + geom_point(data=districthree, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=5)

districtfour<-subset(crime, DISTRICT_ID==4)
str(districtfour)

Denver + geom_point(data=districtfour, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=5)

districtfive<-subset(crime, DISTRICT_ID==5)
str(districtfive)

Denver + geom_point(data=districtfive, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=5)

districtsix<-subset(crime, DISTRICT_ID==6)
str(districtsix)

Denver + geom_point(data=districtsix, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=5)

districtseven<-subset(crime, DISTRICT_ID==7)
str(districtseven)

Denver + geom_point(data=districtseven, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=.5)
table(districtseven$NEIGHBORHOOD_ID)#By dia

Denver + geom_point(data=districtseven, aes(x=GEO_LON, y=GEO_LAT), 
                              color="light green", alpha=.15, size=50)


murders<-subset(crime, OFFENSE_CATEGORY_ID="homicide")

str(murders)
table(murders$DISTRICT_ID)

Denver + geom_point(data=murders, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)
install.packages("rworldmap")             
 library(rworldmap)
 newmap <- getMap(resolution = "low")
plot(newmap, xlim = c(-105.2, -104.8), ylim = c(39.6, 39.9))
points(murders$GEO_LON, murders$GEO_LAT, col = "red", cex = .05)
detach(murders)
summary(crime$OFFENSE_CATEGORY_ID)

murder1<-subset(crime, OFFENSE_CATEGORY_ID="murder")
table(murder1$DISTRICT_ID)

```
Will need to look at the data in different map form
```{r}
rm(list=ls())
crime <-read.csv("C:/Users/BandC/Desktop/Predictive Model/crime.csv")
attach(crime)
summary(crime)
library(ggplot2)
library(ggmap)

murders<-subset(crime, OFFENSE_CATEGORY_ID=="murder")
str(murders)
dim(murders)
library(rworldmap)
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=murders, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)
 newmap <- getMap(resolution = "low")
newmap
plot(newmap, xlim = c(-105.2, -104.8), ylim = c(39.6, 39.9))
points(murders$GEO_LON, murders$GEO_LAT, col = "red", cex = .05)# Not as good as the above map

violent<-subset(crime, OFFENSE_CATEGORY_ID=="murder"|OFFENSE_CATEGORY_ID=="aggravated-assault")
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=violent, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)#can start to see where the high concentrations of violent crime are
# Will want to look at districts too
# Want to look at concentrations by type of crime
type<-subset(crime, DISTRICT_ID==1)
head(type)


Den_con <- get_map(location = 'Denver', source = 'stamen', maptype = 'toner')
ggmap(Den_con) +
  geom_point(size=5, alpha = 1/2, aes(GEO_X,GEO_Y, color=type$OFFENSE_CATEGORY_ID), data=type)

type2<-subset(crime, DISTRICT_ID==2)
head(type2)


Den_con <- get_map(location = 'Denver', source = 'stamen', maptype = 'toner')
ggmap(Den_con) + geom_point(size=5, alpha = 1/2, aes(GEO_X,GEO_Y, color=OFFENSE_CATEGORY_ID), data=type2)
ggmap(Den_con) + stat_bin2d(aes(x = GEO_X, y = GEO_Y, color = OFFENSE_CATEGORY_ID, fill = OFFENSE_CATEGORY_ID), size = 2.5, bins = 30, alpha = 1/2, data = crime) + xlim(c(-105.4,-104.6)) + ylim(c(39.4,40.0))

district1<-subset(crime, DISTRICT_ID==1)
str(district1)
dim(district1)
library(rworldmap)
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=district1, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)

library(ggmap)
district3<-subset(crime, DISTRICT_ID==3)
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=district3, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)
district4<-subset(crime, DISTRICT_ID==4 & OFFENSE_CATEGORY_ID=="aggravated-assault")
str(district4)
dim(district4)
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=district4, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)
district5<-subset(crime, DISTRICT_ID==5 & OFFENSE_CATEGORY_ID=="aggravated-assault")
str(district5)
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=district5, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)
district6<-subset(crime, DISTRICT_ID==6 & OFFENSE_CATEGORY_ID=="aggravated-assault")
str(district6)
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=district6, aes(x=GEO_LON, y=GEO_LAT), 
                              color="red", alpha=.15, size=5)

district7<-subset(crime, DISTRICT_ID==7 & OFFENSE_CATEGORY_ID=="aggravated-assault")
str(district7)
qmap('Denver', zoom=15)
Denver = qmap("Denver", zoom = 11)
Denver + geom_point(data=district7, aes(x=GEO_LON, y=GEO_LAT), 
                              color="blue", alpha=.15, size=5)#Not very concentrated

```
Focusing on crime that is violent. I am going to plot and GIS those instances on a map. First, I am going to get rid of the N/A values with no time or location. I am doing this because I think having a time and locating, 100% accuracy is necessary. Will omit some variables too
```
```{r}
names(crime)
table(crime$OFFENSE_CATEGORY_ID)
finaldata=subset(crime, OFFENSE_CATEGORY_ID=="aggravated-assault" | OFFENSE_CATEGORY_ID=="other-crimes-against-persons" | OFFENSE_CATEGORY_ID=="murder" | OFFENSE_CATEGORY_ID=="robbery" | OFFENSE_CATEGORY_ID=="sexual-assualt")
summary(finaldata)
dim(finaldata)#much less
names(finaldata)
finaldata$IS_CRIME<-NULL #Dropping this variable, since all will be one, since all are violent crime
names(finaldata)
finaldata$IS_TRAFFIC<-NULL

names(finaldata)#Now 17 variables, dropped the two Traffic and Crime, since I am not interested in traffic

pairs(finaldata)#not super revealing but cool 
head(finaldata)
table(finaldata$DISTRICT_ID)#Highest is 3

qplot(finaldata$REPORTED_DATE, binwidth=1)#peak in summer months
days=sort(table(finaldata$REPORTED_DATE),decreasing=TRUE)[1:30]
days

head(finaldata)

finaldata$GEO_X<-NULL

finaldata$GEO_Y<-NULL #These two are duplicated by the GEO_lon and GEO_lat, which are more complete anyways 
names(finaldata)
first_cont=subset(crime, OFFENSE_CATEGORY_ID=="aggravated-assault" | OFFENSE_CATEGORY_ID=="other-crimes-against-persons" | OFFENSE_CATEGORY_ID=="murder" | OFFENSE_CATEGORY_ID=="robbery" | OFFENSE_CATEGORY_ID=="sexual-assualt" & DISTRICT_ID==4)

```

To start looking at models and the data in a better way, I want to do a heirachal clustering. Since it is geographic data, longitude and latitude, I researched some different packages that do clustering with geo distances. Below you will find a few different ways in which I performed clustering

```{r}

library(geosphere)

plot(first_cont$GEO_LON,first_cont$GEO_LAT)
d= as.matrix(dist(cbind(first_cont$GEO_LON,first_cont$GEO_LAT)))
d=as.dist(d)
hc<-hclust(d)
plot(hc)#Shows cluster diagram based on GEO
first_cont$clust <- cutree(hc,k=7)#7 points, for the 7 districts
plot(first_cont$GEO_LON,first_cont$GEO_LAT, col=c("red","blue","green")[first_cont$clust],pch=19)


```
Another approach to Geographic data in clustering

```{r}
install.packages("fields")
library(fields)
dist <- rdist.earth(first_cont,miles = F,R=6371) 
fit <- hclust(as.dist(dist), method = "single")
clusters <- cutree(fit,h = 1000) 
plot(first_cont$GEO_LON, first_cont$GEO_LAT, col = clusters, pch = 20)

```

With a different cut from the original clustering problem

```{r}
library(geosphere)
library(fossil)
d2      <- earth.dist(first_cont)   # distance matrix
hc2     <- hclust(d2)      # hierarchical clustering
plot(hc2)                 # dendrogram suggests 4 clusters
clust <- cutree(hc2,k=4)
clust
install.packages("rgdal")
library(rdgal)
library(ggplot2)
library(ggmap)
map1 <-qmap('Denver', zoom=15)
Denver1 <- qmap("Denver", zoom = 11)
Denver1
denver2<-fortify(Denver1)
denver2
Denver1+
  geom_path(aes(x=first_cont$GEO_LON, y=first_cont$GEO_LAT))+
  geom_point(data=first_cont, aes(x=GEO_LON, y=GEO_LAT, color=factor(clust)), size=4)
  scale_color_discrete("Cluster")+
  coord_fixed()
d3      <- earth.dist(first_cont)   # distance matrix
hc3     <- hclust(d2)      # hierarchical clustering
plot(hc3)                 # dendrogram suggests 4 clusters
clust2 <- cutree(hc3,k=7)
clust2
Denver1+
  geom_path(aes(x=first_cont$GEO_LON, y=first_cont$GEO_LAT))+
  geom_point(data=first_cont, aes(x=GEO_LON, y=GEO_LAT, color=factor(clust2)), size=7)+
  scale_color_discrete("Cluster")+
  coord_fixed()

```
With just murder and aggravated assault

```{r}
datafinal=subset(finaldata, OFFENSE_CATEGORY_ID=="aggravated-assault" | OFFENSE_CATEGORY_ID=="murders")
na.omit(datafinal)
library(geosphere)
head(datafinal)

plot(datafinal$GEO_LON,datafinal$GEO_LAT)
dff= as.matrix(dist(cbind(datafinal$GEO_LON,datafinal$GEO_LAT)))
dff=as.dist(dff)
hcf<-hclust(dff)
plot(hcf)#Shows cluster diagram 
datafinal$clust <- cutree(hcf,k=7)#7 points, for the 7 districts
plot(finaldata$GEO_LON,datafinal$GEO_LAT, col=c("red","blue","green")[datafinal$clust],pch=19)
library(fields)
distf <- rdist.earth(datafinal,miles = F,R=6371) 
fitf <- hclust(as.dist(distf), method = "single")
clustersf <- cutree(fitf,h = 1000) 
plot(datafinal$GEO_LON, datafinal$GEO_LAT, col = clusters, pch = 20)
library(geosphere)
library(fossil)
df      <- earth.dist(datafinal)   # distance matrix
hcf     <- hclust(df)      # hierarchical clustering
plot(hcf)                 # dendrogram suggests 4 clusters
clustf <- cutree(hcf,k=4)
clustf

library(rdgal)
library(ggplot2)
library(ggmap)
map1 <-qmap('Denver', zoom=15)
Denverf <- qmap("Denver", zoom = 11)
Denverf
library(ggplot2)
denver2f<-fortify(Denverf)
denver2f
Denverf+
  geom_path(aes(x=datafinal$GEO_LON, y=datafinal$GEO_LAT))+
  geom_point(data=datafinal, aes(x=GEO_LON, y=GEO_LAT, color=factor(clustf)), size=4)+
  scale_color_discrete("Cluster")+
  coord_fixed()
df3      <- earth.dist(datafinal)   # distance matrix
hcf3     <- hclust(df3)      # hierarchical clustering
plot(hc3)                 # dendrogram suggests 4 clusters
clustf2 <- cutree(hcf3,k=7)
clustf2
Denver1+
  geom_path(aes(x=datafinal$GEO_LON, y=datafinal$GEO_LAT))+
  geom_point(data=datafinal, aes(x=GEO_LON, y=GEO_LAT, color=factor(clustf2)), size=7)+
  scale_color_discrete("Cluster")+
  coord_fixed()

```
In addition to clustering,I want to do another modeling technique. I will use a tree.

```{r}
murder<-subset(finaldata, OFFENSE_CATEGORY_ID=="murder")
head(murder)
na.omit(murder)
dim(murder)
head(murder)
install.packages("randomForest")
library(randomForest)
names(finaldata)
train=sample(1:nrow(murder), nrow(murder)/2)
forest=randomForest(OFFENSE_CODE~ OFFENSE_CATEGORY_ID + DISTRICT_ID + PRECINCT_ID + GEO_LON + GEO_LAT,data=murder, subset=train, mtry=15, importance= TRUE)
forest
murdertest=murder[-train,"OFFENSE_CODE"]
y=predict(forest, newdata=murder[-train,])
plot(y,murdertest)
mean((y-murdertest)^2)# This does not really do much, will want to do classification for any really good results
#Let's see if we can classify murder by the district I live in
library(tree)


murderclass=ifelse(DISTRICT_ID=="4","No","Yes")
start=data.frame(murder,murderclass)
head(murder)
head(start)
summary(start)
treestart=tree(murderclass~., start)
summary(treestart)
plot(treestart)
text(treestart)
treestart
dim(start)
train=sample(1:nrow(start))
treestart.test=start[-train,]
id=start[-train]
id
tree1=tree(murderclass~.-DISTRICT_ID, start,subset=train)
tree.pred=predict(tree1,id,type="class")
table(tree.pred,id)
#.74#It looks like this is the percentage correct
rm(list=ls())
#For a prediction, let's look at logistic
crime2 <-read.csv("C:/Users/BandC/Desktop/Predictive Model/crime2.csv")

                  head(crime2)
na.omit(crime2)
dim(crime2)
str(crime2)
crime2$GEO_X<-NULL
crime2$GEO_Y<-NULL
crime2$LAST_OCCURENCE_DATE<-NULL
crime2$FIRST_OCCURENCE_DATE<-NULL
crime2$IS_TRAFFIC<-NULL
crime3<-subset(crime2,crime2$DISTRICT_ID==4 & NEIGHBORHOOD_ID=="athmar-park")
head(crime3)
crime3$IS_CRIME<-as.factor(crime3$IS_CRIME)
str(crime3)
dim(crime3)
attach(crime3)
train_ind = sample(nrow(crime3), nrow(crime3)/2) 
train_data = crime3[train_ind,]
dim(train_data)
test_data = crime3[-train_ind,]
dim(test_data)

yes = glm(IS_CRIME~ OFFENSE_CATEGORY_ID + REPORTED_DATE + OFFENSE_CATEGORY_ID, data=train_data, family=binomial(link='logit'))
summary(yes)
predict = predict(yes, type="response")
predict[1:10]
contrasts(crime3$IS_CRIME)
glm.pred=rep("no",1577)
glm.pred[predict>.5]="yes"
table(glm.pred,train_data$IS_CRIME)##glm.pred   no  yes, not great 
     no   343    4
     yes    2 2428




```

Based on the analsis, the clustering was very productive as it used longitude and latitude to cluster. This approach with using geographical distances was valuable. The logistic approach did not seem to work well and was not the best model. 
```

